---
title: "Categorizing posts"
output: github_document
---

## Defining categories

I reviewed the list of most-common words in the ecolog subject headers, and came up with a classification system for the different types of posts:

- positions
    - jobs
    - internships
    - volunteer
    - reu
    - msc
    - phd
    - postdoc
    - faculty
- events
- grants/scholarships

Below are some words that belong to each category, as well as synonyms. *Note:* I've included only the singular/root words for families of terms (e.g. intern = interns/internship).

#### Jobs
job = hiring = employment

temporary
intern
volunteer

scientist
ecologist
researcher
technician
lead
assistant
fellow
educator
specialist
manager
biologist
coordinator
director

#### Academic positions
reu = undergrad

msc = m.sc = master = graduate

phd = ph.d = doctoral = graduate

postdoc = postdoctoral

professor = faculty = tenure

lecturer = instructor

#### Events
program
project
conference
workshop
training
course
webinar

#### Funding
grant
scholarship
proposal
funding

## Implementation

Now let's define some variables and functions to help with the categorization process. We'll start with the two easier categories using a simple `str_detect`:

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(here)

is_event <- function(subject) {
    stringr::str_detect(subject, "conference|workshop|training|course|webinar")
}

is_funding <- function(subject) {
    stringr::str_detect(subject, "grant|scholarship|proposasl|funding")
}
```



```{r}
job_synonyms <- "job hiring employment"

position_names <-  "scientist ecologist researcher technician lead assistant fellow educator specialist manager biologist coordinator director"

aca_positions <- c(undergrad = "reu undergrad",
                   masters = "msc m.sc master graduate",
                   phd = "phd ph.d doctoral graduate",
                   postdoc = "postdoc postdoctoral",
                   faculty = "professor faculty tenure lecturer instructor")

misc_positions <- c("temporary", "intern", "volunteer")

all_positions <- paste(job_synonyms, position_names, 
                       paste(aca_positions, collapse = " "), 
                       paste(misc_positions, collapse = " "))

is_position <- function(subject){
    str_detect(subject, 
               all_positions %>% str_replace_all(" ", "|"))
}
```

Now let's try it out on our data!

```{r}
ecolog_2018_df <- read.csv(here("data", "ecolog_2018.csv")) #do this outside of rmd

ecolog_2018_df %>% 
    mutate(funding = is_funding(subject),
           event = is_event(subject),
           position = is_position(subject)) %>% 
    head()
```

